---
title: "p8105_hw2_qz2392"
author: "Qimin Zhang"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Problem 1
Read the data in Mr.Trash Wheel sheet without rows and columns with notes, and build pipeline for cleaning tha data.
```{r}
mrtrashwheel = 
  read_excel(
    "./Datasets/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    range = "A2:N406") %>%
  janitor::clean_names() %>%
  drop_na("dumpster") %>%
  mutate(
    sports_balls = as.integer(round(sports_balls))
    )
```

Read and clean the data in 2017 and 2018 Precipitation sheet.
```{r}
#Read and clean the 2017 Precipitation sheet
pre_2017 =
  read_excel(
    "./Datasets/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", 
    range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na("total") %>%
  mutate(
    year = 2017
    )

#Read and clean the 2018 Precipitation sheet
pre_2018 =
  read_excel(
    "./Datasets/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation", 
    range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na("total") %>%
  mutate(
    year = 2018
    )

#Combine the precipitation sheets and convert the month variable into character type.
pre_2017_and_2018 =
  bind_rows(pre_2017, pre_2018) %>%
  janitor::clean_names() %>%
  mutate(
    month = month.name[month]
  )

```
The number of observations in Mr.Trash Wheel dataset is `r nrow(mrtrashwheel)` and the number of observations in precipitation data for 2017 and 2018 is `r nrow(pre_2017_and_2018)`. The key variables in Mr.Trash Wheel dataset might include 'weight_tons' and 'homes_powered', and the key variables in precipitation data for 2017 and 2018 might include 'month' and 'total'.
The total precipitation in 2018 is `r sum(select(pre_2018, total))`, and the median number of sports balls in a dumpster in 2017 is `r median(pull(filter(mrtrashwheel, year == 2017), sports_balls))`.

# Problem 2
First to clean the dataset pols-month.
```{r, message=FALSE}
pols_month =
  read_csv("./Datasets/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, c('year', 'month', 'day'), sep = '-') %>%
  mutate(
    year = as.integer(year),
    month = month.name[as.integer(month)],
    day = as.integer(day),
    president = case_when(
  prez_gop > 0 ~ 'gop',
  prez_dem > 0 ~ 'dem'
)
  ) %>%
  select(-c(day,prez_dem, prez_gop))
```
Then to clean the dataset snp.
```{r, message=FALSE}
snp =
  read_csv("./Datasets/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, c('month', 'day', 'year'), sep = '/') %>%
  mutate(
    year = as.integer(year),
    month = month.name[as.integer(month)],
    day = as.integer(day)
  ) %>%
  select(year, month, close, -day)
```
Next to tidy the unemployment dataset so that it can be merged with the previous datasets.
```{r, message=FALSE}
unemployment =
  read_csv("./Datasets/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>%
  janitor::clean_names() %>%
  mutate(
    month = case_when(
      month == 'Jan' ~ 'January',
      month == 'Feb' ~ 'February',
      month == 'Mar' ~ 'March',
      month == 'Apr' ~ 'April',
      month == 'May' ~ 'May',
      month == 'Jun' ~ 'June',
      month == 'Jul' ~ 'July',
      month == 'Aug' ~ 'August',
      month == 'Sep' ~ 'September',
      month == 'Oct' ~ 'October',
      month == 'Nov' ~ 'November',
      month == 'Dec' ~ 'December'
    ) 
  )
```
Now join the datasets.
```{r}
fivethirtyeight =
  left_join(left_join(pols_month, snp, by = c("year", "month")), 
            unemployment, 
            by = c("year", "month"))
```
The pol_month dataset contains the information about the number of govenors, senators and representitves from republican and democratic parties, and specifies the party the president belonged to on the associated date. The snp dataset contains the closing values of the S&P stock index on the associated date. The unemploymeny dataset contains the percentage of unemployment on the associated date.
The dimension of the resulting dataset is `r nrow(fivethirtyeight)` rows and `r ncol(fivethirtyeight)` columns. The range of years is from `r min(pull(fivethirtyeight, year))` to `r max(pull(fivethirtyeight, year))` years. The names of key variables are 'president', 'rep_gop', 'rep_dem' and 'unemployment_rate', because the president and the ratio of republican govenors to democratic govenors can reflect to the power of a specific party, and the unemployment rate can reflect the economy status.

# Problem 3
Read and clean the popular baby names dataset.
```{r, message=FALSE}
popular_baby_names =
  read_csv("./Datasets/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    gender = str_to_upper(gender),
    ethnicity = str_to_upper(ethnicity),
    ethnicity = case_when(
      ethnicity == "ASIAN AND PACIFIC ISLANDER" ~ "ASIAN AND PACIFIC ISLANDER",
      ethnicity == "BLACK NON HISPANIC" ~ "BLACK NON HISPANIC",
      ethnicity == "WHITE NON HISPANIC" ~ "WHITE NON HISPANIC",
      ethnicity == "ASIAN AND PACI" ~ "ASIAN AND PACIFIC ISLANDER",
      ethnicity == "BLACK NON HISP" ~ "BLACK NON HISPANIC",
      ethnicity == "WHITE NON HISP" ~ "WHITE NON HISPANIC",
      ethnicity == "HISPANIC" ~ "HISPANIC"
    ),
    childs_first_name = str_to_title(childs_first_name)
  ) %>%
  distinct()
```
Investigate the popularity of the name 'Olivia' over time among female babies.
```{r, message= FALSE, warning = FALSE}
olivia_rank =
  popular_baby_names %>%
  filter(childs_first_name == "Olivia" & gender == "FEMALE") %>%
  select(-c(gender, childs_first_name, count)) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>%
  knitr::kable()

olivia_rank
```

Investigate the most popular name over time among male babies.
```{r}
male_rank =
  popular_baby_names %>%
  filter(gender == "MALE" & rank == 1) %>%
  select(-c(gender, rank, count)) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>%
  knitr::kable()
male_rank
```

For male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).
```{r, message = FALSE}
scatterplot_names_poplarity =
  popular_baby_names %>%
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) %>%
  ggplot(aes(x = rank, y = count)) +
  geom_point()+
  labs(
    title = "Baby names popularity",
    x = "Baby names rank",
    y = "Baby names count"
  ) +
  theme_minimal()

scatterplot_names_poplarity
ggsave("scatterplot_names_poplarity.pdf", scatterplot_names_poplarity)
```