---
title: "p8105_hw2_qz2392"
author: "Qimin Zhang"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1
Read the data in Mr.Trash Wheel sheet without rows and columns with notes, and build pipeline for cleaning tha data.
```{r}
mrtrashwheel = 
  read_excel(
    "./Datasets/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    range = "A2:N338") %>%
  janitor::clean_names() %>%
  drop_na("dumpster") %>%
  mutate(
    sports_balls = as.integer(round(sports_balls))
    )
```

Read and clean the data in 2017 and 2018 Precipitation sheet.
```{r}
#Read and clean the 2017 Precipitation sheet
pre_2017 =
  read_excel(
    "./Datasets/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", 
    range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na("total") %>%
  mutate(
    year = 2017
    )

#Read and clean the 2018 Precipitation sheet
pre_2018 =
  read_excel(
    "./Datasets/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet = "2018 Precipitation", 
    range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na("total") %>%
  mutate(
    year = 2018
    )

#Combine the precipitation sheets and convert the month variable into character type.
pre_2017_and_2018 =
  bind_rows(pre_2017, pre_2018) %>%
  janitor::clean_names() %>%
  mutate(
    month = month.name[month]
  )

```
Paragraph. 
The total precipitation in 2018 is `r sum(select(pre_2018, total))`, and the median number of sports balls in a dumpster in 2017 is `r median(pull(filter(mrtrashwheel, year == 2017), sports_balls))`.

# Problem 2
First to clean the dataset pols-month.
```{r, message=FALSE}
pols_month =
  read_csv("./Datasets/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, c('year', 'month', 'day'), sep = '-') %>%
  mutate(
    year = as.integer(year),
    month = month.name[as.integer(month)],
    day = as.integer(day),
    president = case_when(
  prez_gop > 0 ~ 'gop',
  prez_dem > 0 ~ 'dem'
)
  ) %>%
  select(-c(day,prez_dem, prez_gop))
```
Then to clean the dataset snp.
```{r, message=FALSE}
snp =
  read_csv("./Datasets/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, c('month', 'day', 'year'), sep = '/') %>%
  mutate(
    year = as.integer(year),
    month = month.name[as.integer(month)],
    day = as.integer(day)
  ) %>%
  select(year, month, close, -day)
```
Next to tidy the unemployment dataset so that it can be merged with the previous datasets.
```{r, message=FALSE}
unemployment =
  read_csv("./Datasets/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>%
  janitor::clean_names() %>%
  mutate(
    month = case_when(
      month == 'Jan' ~ 'January',
      month == 'Feb' ~ 'February',
      month == 'Mar' ~ 'March',
      month == 'Apr' ~ 'April',
      month == 'May' ~ 'May',
      month == 'Jun' ~ 'June',
      month == 'Jul' ~ 'July',
      month == 'Aug' ~ 'August',
      month == 'Sep' ~ 'September',
      month == 'Oct' ~ 'October',
      month == 'Nov' ~ 'November',
      month == 'Dec' ~ 'December'
    ) 
  )
```
Now join the datasets.
```{r}
fivethirtyeight =
  left_join(left_join(pols_month, snp, by = c("year", "month")), 
            unemployment, 
            by = c("year", "month"))
```
Paragraph. 
The pol_month dataset contains the information about the number of govenors, senators and representitves from republican and democratic parties, and specifies the party the president belonged to on the associated date. The snp dataset contains the closing values of the S&P stock index on the associated date. The unemploymeny dataset contains the percentage of unemployment on the associated date.

# Problem 3
Read and clean the popular baby names dataset.
```{r, message=FALSE}
popular_baby_names =
  read_csv("./Datasets/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  distinct()
```
Investigate the popularity of the name 'Olivia' over time among female babies.
```{r, message= FALSE, warning = FALSE}
olivia_rank =
  popular_baby_names %>%
  filter(childs_first_name == "Olivia" & gender == "FEMALE") %>%
  select(-c(gender, childs_first_name, count)) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )

olivia_rank
```
Investigate the most popular name over time among male babies.
```{r}
male_rank =
  popular_baby_names %>%
  filter(gender == "MALE" & rank == 1) %>%
  select(-c(gender, rank, count)) %>%
  mutate(
    ethnicity = case_when(
      ethnicity == "ASIAN AND PACIFIC ISLANDER" ~ "ASIAN AND PACIFIC ISLANDER",
      ethnicity == "BLACK NON HISPANIC" ~ "BLACK NON HISPANIC",
      ethnicity == "WHITE NON HISPANIC" ~ "WHITE NON HISPANIC",
      ethnicity == "ASIAN AND PACI" ~ "ASIAN AND PACIFIC ISLANDER",
      ethnicity == "BLACK NON HISP" ~ "BLACK NON HISPANIC",
      ethnicity == "WHITE NON HISP" ~ "WHITE NON HISPANIC",
      ethnicity == "HISPANIC" ~ "HISPANIC",
    )
  ) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )

male_rank
```


